Using reference models in feature selection problems \vspace{.1in}

### 2019-07-30 paper/ref_approach.tex
start of the re-organization interrupted by vacation

\begin{document}
\maketitle
\begin{abstract}
Feature selection, or more generally, model reduction is an important aspect of statistical workflow aiming to provide insights from data. While a vast number of feature selection procedures are available, it often remains unclear what are the underlying principles explaining why certain methods are working better than others. We investigate and discuss one of these principles in detail, which we identify with the use of a reference model. We demonstrate how reference models act as a noise-filters on the target variable by approximating its data generation mechanism.
To this end, we devise a simple reference model approach that can be used on top of any feature selection method. In several numerical experiments, we show how this translates into better and more stable feature selection independently of the specific selection procedure. 
These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.

%When a reference model is used to guide the variable selection problem, it acts as a noise-filter modelling the data generation mechanism. We show how this translate into higher stability and better selection of variables. We devise a simple reference model approach that can be used on top of any feature selection procedure and use the normal means problem as a benchmark comparing different methods using or not a reference model. We include in our comparisons state-of-the-art complete selection procedures showing improved stability, false discovery rate and sensitivity of the selection, and also Bayesian methods controlling the amount of posterior shrinkage and bias. Furthermore we show in a real-world example the benefits achieved by the reference model through the projection predictive approach compared to the selection via stepwise backward regression.

\end{abstract}

@@

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all covariates. The reference model approach has been
used for long time in Bayesian statistics since the seminal work of
\citet{paper:reference_lindley}. For more historical references, see
papers by \citet{vehtari2012survey} and \citet{paper:model_selection},
and for most recent methodological development see
\citet{paper:projpred}.

@@

In this paper, we examine the benefits of using reference models in
detail. We illustrate how a properly designed reference model is able
to filter parts of the noise present in the data, and hence provide an
improved and more stable selection process. This holds even if the
reference model does not perfectly resemble the true data generating
process. Moreover, our analyses indicate that the substantial
reduction of variance attributable to noise is usually more important
than small potential bias due to model misspecification.  We argue
that, regardless of how the reference model is set up and used in the
inference procedure, it can be always seen as acting as a filter on
the observed data. Following this idea, we devise a simple reference
model approach that can be applied on top of any variable selection
procedure.

In our simulations and analysis of real data, we apply several
state-of-the-art methods for variable selection and compare, for each
of these methods, the results obtained by using a data based and a
reference model based approach. When a full variable selection is
carried out, we report and compare quantities such as the false
discovery rate \citep{paper:fdr_BH}, sensitivity and stability of the
selection \citep{paper:stability}. For fully Bayesian methods, we
additionally report results in terms of shrinkage and accuracy of the
posterior estimates.  We show the benefits of a reference model
approach, regardless of what specific procedure is applied, using the
normal means estimation problem as a benchmark.  Variable selection is
typically most challenging in \textit{small n large p} scenarios, that
is, scenarios with small number of observations but a lot of unknown
model parameters.  These occur frequently, for example, in microarray
data analysis in medicine or biology
\citep{dudoit2003multiple,liao2007logistic,paper:efron}.  In general,
the less data we have and the more complex the estimated models, the
higher the benefit of using a reference models as the basis for
variable selection.  Our results indicate that the core reason why the
reference model based methods perform well is the reference model
itself, rather than the specific way of using it. We do recognise that
all these methods diversify how the subset of variables is actually
chosen and which properties it has, thus it remains fundamental to
investigate and develop such different procedures. Although we focus
on Bayesian approach in this paper, the benefits of the reference
model approach can also be obtained in non-Bayesian settings.

The remainder of the paper is structured as follows. In Section
\ref{reference-model-approach}, we review the concept of the reference
model, its benefits with examples, including the projection predictive
approach, and how it can be used as a filter on data in a simple
way. In Section \ref{comparison}, we use different methods for feature
selection comparing the selection results with and without using a
reference model in the framework of the normal means problem. Finally,
we end with a conclusion in Section \ref{conclusion}.

@@

The main assumption under any reference model approach is that we
operate in an $\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is we have constructed a
model which reflects in the best way our beliefs about the future
data. In practice, this would often be the full-encompassing model
(including all models we consider having non-zero prior
probability as submodels) which has passed model checking and criticism \citep[see,
e.g.][]{gelman2013bayesian}.

A good predictive model is able to filter part of the noise present in
the data. The noise is the main source of the instability and tends to
obscure the relevance of the covariates in relation to the target
variable of interest. Indeed, let us consider the following
explanatory example taken from \cite{paper:projpred}. The data
generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim N(0,1) && \nonumber \\ 
     Y|&f\sim N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} N(\sqrt{\rho}f,1-\rho) \quad &&j=1,..,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} N(0,1) &&j=k+1,..,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy
observation. The first $k$ covariates are strongly related to the
target variable $Y$ and correlated among themselves. Precisely, $\rho$
is the correlation among any pair of the first $k$ covariates, whereas
$\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between
any relevant covariate and, respectively, $f$ and $Y$. If we had an
infinite amount of observations, the sample correlation would be equal
to the true correlation between $X_j$ and $Y$. However, even in this
ideal asymptotic regime, this correlation would still remain a biased
indicator of the true relevance of each covariate (represented by the
correlation between $X_j$ and $f$) due to the intrinsic noisy nature
of $Y$.

@@

\hypertarget{projection}{%
\subsection{Benefits of the reference model: an example with the projection predictive approach}\label{projection}}

\subsubsection{Example: Variable selection in bodyfat prediction}

We show the benefits of the reference model in the form of the
projection approach (projpred) with respect to the stepwise backward
regression (steplm) using the body fat data
\citep{johnson1996fitting}\footnote{The analysis is based on
  \texttt{R}-notebook by Vehtari available at
  \url{https://avehtari.github.io/modelselection/bodyfat.html}.}. The
target variable of interest is the amount of body fat, which is
measured with a complex and expensive procedure. Additionally, we have
information about 13 covariates which are anthropometric measurements
(e.g., height and weight), some of which are highly correlated among
themselves and thus provide some challenges in the model selection. In
total, we have 251 observations. The goal is to find the simplest
model which is able to predict the amount of fat sufficiently well. We
use the same data again later on in Section \ref{real-world-data}.

\cite{paper:bodyfat} report results using steplm with a significance
level of 0.157 with AIC selection \citep{akaike1974new}, fixing
abdomen and height to be always included in the model. Here, we
implement steplm via AIC without fixing any variable. For the
selection via projpred, the reference model includes all the
covariates, we regularise model complexity with a regularised
horseshoe prior on the covariate coefficients \citep{paper:rhs},
submodels are explored using forward search, and the predictive
utility is the expected log-predictive density (elpd) estimated using
PSIS-LOO \citep{paper:psis_loo}. The experiments are implemented in
\texttt{R} \citep{Rcore2018}. The steplm approach is carried out
combining the \texttt{step} and \texttt{lm} functions, whereas the
selection via projpred is done with the \texttt{projpred} package. In
order to automatise the latter, we select the submodel as the one with
the smallest size which has an elpd score higher (i.e., better) than
the reference model score with probability of at least 5\%. The
inclusion frequencies of each feature in the final model after 100
bootstrap samples are depicted in Figure
\ref{fig:inclusion_frequencies}. The projpred approach has only two
variables, `abdomen' and `weight', with inclusion frequencies above
50\% (`abdomen' is the only one included always), the third most
included is `wrist' at 44\%, and the fourth is `height' at 35\%, while
steplm has seven features above 50\%. Such a higher variability and
lower stability of steplm can be observed also in the bootstrap model
selection frequencies reported in Table
\ref{tab:model_frequencies}. For example, the first five selected
models have a cumulative frequency of 76\% with projpred, but only of
14\% with steplm. In addition, the size of the selected model is much
smaller when using projpred.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/inc_prob.pdf}
  \caption{Bootstrap inclusion frequencies after 100 bootstrap samples.}
  \label{fig:inclusion_frequencies}
\end{figure}

First two rows of Table \ref{tab:model_performances} report the
predictive performances, in terms of root mean square error (RMSE), of
the full model and the selected models using either the projection
predictive approach or the stepwise backward elimination.  There is no
difference in predictive performance of the selected models by
different approaches even if there is difference in the selected
features.

We repeat the experiment with a modified data set by adding 84
unrelated noisy features. Last two rows of Table
\ref{tab:model_performances} report RMSE, the size of the selected
model and the number of selected noisy features using either the
projection predictive approach or the stepwise backward elimination.
The projection predictive approach has similar predictive performance
and the same submodel size as in the first experiment, whereas the
stepwise regression has worse predictive performance and selects a
much larger submodel including a large number of irrelevant features.

\hypertarget{computational-burden}{%
\subsection{Computational cost} \label{computational-burden}}

There are various ways to implement a reference model approach. A
common term in the computational cost is the inference for the
reference model. In the simple ``bare-bones'' reference model approach
we have described above, it is the only added cost. In general, a full
Bayesian reference model is advisable in order to have the best
predictive performance. However, including all the available features
can often be quite computationally demanding. In such a case, a first
possible speed-up could be using screening or dimensionality reduction
techniques, as for example the supervised principal components
\citep{bair2006prediction,piironen2018}. However, in some cases, this
does not always result in an appreciable speed-up. For example, when
there are several predictive but mutually uncorrelated features, all
of them have to be included in the reference model. Other alternatives
could be using point estimates such maximum a posteriori estimatge if
predictions from such reference model without full Bayesian inference
are good enough.

The comparison example between projpred and steplm shows that such
additional computational cost can be worth it. We found that the
reference model approach (projpred) selected much sparser submodels,
which can result in improved and more stable interpretability and
saving time and general costs using the selected model for future
predictions. Table \ref{tab:model_performances} and results by
\citet{paper:model_selection} demonstrate also that in case of large
number of features, there can be a significant improvement in the
predictive ability compared to feature selection without a reference
model, thus justifying additional computational costs in the selection
procedure. More benefits obtainable through a reference model approach
are provided in the experiments discussed in the upcoming section.

@@

In this section, we analyse in more detail the benefits of reference
model approaches using the normal means problem as a benchmark. The
normal means problem consists of estimating the (usually sparse)
vector of means of a vector of normally distributed observations. The
dimensionality of the vector of means is denoted by $p$ and
$\{z_{j}\}_{j=1}^{p}$ is the vector of observations of the random
variables $\{Z_{j}\}_{j=1}^{p}$. In the normal means problem, the task
is to estimate the latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the
following model: \
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}), \quad j=1,..,p.
\end{equation}
This is equivalent to a linear regression where the design matrix is
the identity matrix with the number of features being equal to the
number of observations. This formulation can be found in practice, for
example, in the analysis of microarray data, where a large set of
genes are tested in two groups of patients labeled as positive or
negative to some disease \citep{paper:efron,efron2012large}. The
objective of the analysis is to select the subset of genes
statistically relevant to the disease. One common way to proceed is to
compute the two-sample $t$-statistic for every gene separately.  After
normalising these statistics, they then become the data, $Z$, in the
normal means problem \eqref{eq:normal_means_problem}. For further
details see the examples in \cite{paper:efron, efron2012large}.

In our experiments, we retrieve the normal means problem from the
sample correlations between the target and the covariates using the
Fisher $z$-transformation \citep{hawkins1989using}. Suppose we have a
continuous target random variable $Y$, a set of $p$ continuous
covariates $\{X_{j}\}_{j=1}^{p}$, and denote
$\rho_{j}=Cor(Y,X_{j})$. Further, suppose we have observed $n$
statistical units and define $r_{j}$ the sample correlation between
the observations of the target variables $\{y_{i}\}_{i=1}^{n}$ and the
$j$-th covariate $\{x_{ij}\}_{i=1}^{n}$. Finally, we refer to the
Fisher $z$-transformation function $\text{tanh}^{-1}(\cdot)$ as
$T_{F}(\cdot)$. Assuming each pair $(Y,X_{j})$ to be bivariate
normally distributed, the corresponding transformed correlations are
approximately normally distributed with known variance: \
\begin{equation} \label{eq:fisher_transformation}
T_{F}(r_{j})\overset{ind}{\sim} N(T_{F}(\rho_{j}),\frac{1}{n-3}) \quad j=1,..,p.
\end{equation}
Therefore, rescaling the quantities $T_{F}(r_{j})$ by $\sqrt{n-3}$ and
denoting the results as $z_{j}$, we have the formulation
\eqref{eq:normal_means_problem} of the normal means problem, this time
with unit variance: \
\begin{equation} \label{eq:normal_means_problem2}
Z_{j}|\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \quad j=1,..,p.
\end{equation}
Now the quantities of interest $\theta_{j}$ are equal to
$\sqrt{n-3}\,T_{F}(\rho_{j})$. There are different ways to proceed
with the selection in the normal means problem. Here, we consider
several state-of-the-art approaches primarily from Bayesian
statistics.

We briefly review some Bayesian estimation methods to solving the
normal means problem. They all have in common that a Bayesian
regression model is fitted using some sparsifying prior on the
regression coefficients, for example, as discussed in
\cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and
\cite{johnstone2004needles}. If these priors have continuous support,
as is the case for the Dirichlet-Laplace
\cite[DL;][]{paper:dirichlet_laplace} or horseshoe-like priors
\citep{paper:hs,bhadra2017horseshoe+,paper:rhs}, there is no canonical
way to complete the selection since the posterior probability of an
exactly zero signal is always zero. Whereas
\cite{paper:dirichlet_laplace} used k-means clustering over the
posterior samples to perform the actual selection,
\cite{bhadra2017horseshoe+} just focused their analysis on the
obtained shrinkage and the correct estimation of the sparse signals
without any actual selection. In our experiments, we include the DL
prior and the regularised horseshoe (RHS) prior \citep{paper:rhs},
which we think scales better to higher number of covariates as
compared to other horseshoe-like priors. For the Bayesian regression
models with these priors, we compare the posterior shrinkage and the
correct estimation of the sparse signals with and without using a
reference model as a noise filter. For this purpose, we consider the
average SSE (sum of square errors) and the average SESE (sum of
expected square errors) as quantities measuring estimation
accuracy. They are defined as \
\begin{align}
\text{SSE}&=\frac{1}{n-3}\sum_{j=1}^{p}(\hat{\theta}_{j} - \theta^{0}_{j})^{2} \label{eq:SSE} \\
\text{SESE}&=\frac{1}{n-3}\sum_{j=1}^{p}\int_{\mathbb{R}}(\theta_{j}-\theta^{0}_{j})^{2}p(\theta_{j}|z_{1},..,z_{p})d\theta_{j} \\
&=\frac{1}{n-3}\sum_{j=1}^{p}\mathbb{E}_{\theta_{j}}[(\theta_{j}-\theta^{0}_{j})^{2}|z_{1},..,z_{p}], \label{eq:SESE}
\end{align}
where $\theta_{j}^{0}=\sqrt{n-3}\,T_{F}(\rho_{j})$ is the true value,
$\hat{\theta}_{j}$ is the posterior median of each parameter
$\theta_{j}$. We divide the errors by $(n-3)$ to have quantities
independent of the number of statistical units.  The SESE takes the
posterior uncertainty in account and thus also the shrinkage of the
posterior distributions, whereas the SSE only measures the bias in the
point estimates.  Any (sensible) point estimate could be used to
compute the SSE but here we follow \cite{paper:dirichlet_laplace} and
use the posterior median. The results that we show later on in Section
\ref{shrinkage-signal} are not sensitive to the choice of this point
estimate.

For shrinkage priors, one central aspect is the amount of shrinkage
they actually impose on the relevant and irrelevant features. Ideally,
the irrelevant features should be completely shrunk to zero while the
relevant features should not be shrunk at all. The amount of shrinkage
can be quantified by the shrinkage factor $\kappa_{j}$ for each
parameter $\theta_{j}$ ranging from $0$ (no shrinkage) to $1$
(complete shrinkage). For model \eqref{eq:normal_means_problem2} with
a hierarchical shrinkage prior of the type
\begin{align}
Z_{j}|&\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \qquad j=1,..,p \\
&\theta_{j}|\tau,\nu_{j}\overset{ind}{\sim}N(0,\nu_{j}^{2}\tau^{2})
\end{align}
the shrinkage factor $\kappa_{j}$ is defined as
\begin{equation}
\kappa_{j}=\frac{1}{1+\tau^{2}\nu_{j}^{2}}.
\end{equation}
\
For RHS and DL priors we have
\begin{equation}
\kappa_{j}^{\text{RHS}}=\frac{1}{1+\tau^{2}\lambda_{j}^{2}}, \qquad \kappa_{j}^{\text{DL}}=\frac{1}{1+\psi\phi_{j}^{2}\tau^{2}}.
\end{equation}
In the left hand-side expression, $\tau$ and $\lambda_{j}$ represent
the global and local shrinkage parameters of the regularised horseshoe
prior \cite[for further details see][]{paper:rhs}, whereas on the
right hand-side, $\psi\phi_{j}^{2}\tau^{2}$ is the variance term of
the Dirichlet-Laplace prior of parameter $\theta_{j}$ following the
notation of \cite{paper:dirichlet_laplace}. We use the posterior mean
of $\kappa_{j}$ as a point estimate.

>>>>

\begin{document}
\maketitle
\begin{abstract}
  Feature selection, or more generally, model reduction is an important aspect of statistical workflow aiming to provide insights from data. In this paper we discuss and demonstrate the benefits of using a reference model in model selection in general. A reference model acts as a noise-filter on the target variable by modeling its data generation mechanism. Using the reference model predictions in the model selection reduces tha variability and improves stability leading to improved model selection performance. The simplest reference model approach can be used in combination of common feature selection methods . In several numerical experiments, we show how this translates into better and more stable feature selection independently of the specific selection procedure. 
% These benefits are also illustrated on a real-world example of predicting the amount of body fat for which we compare a reference model based selection method to a purely data driven approach.
\end{abstract}

@@

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all covariates.  The main assumption under any
reference model approach is that we operate in an
$\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is we have
constructed a model which reflects in the best way our beliefs about
the future data and which has passed model checking and criticism
\citep[see, e.g.][]{gelman2013bayesian}.  The reference model approach
has been used for long time in Bayesian statistics since the seminal
work of \citet{paper:reference_lindley}. For more historical
references, see papers by \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for most recent methodological
development see \citet{paper:projpred}.

@@

\subsection{Example of model selection with or without reference model}

To motivate the further discussion and experiments, we start by simple
example comparing projection predictive approach (projpred,
\citet{paper:projpred}) which uses a reference model and classic
stepwise backward regression (steplm).  using the body fat data
\citep{johnson1996fitting}\footnote{The analysis is based on
  \texttt{R}-notebook by Vehtari available at
  \url{https://avehtari.github.io/modelselection/bodyfat.html}.}.  The
experiments are implemented in \texttt{R} \citep{Rcore2018}.

The target variable of interest is the amount of body fat, which is
measured by immersing a person in water tank which is a complex and
expensive procedure. Additionally, we have information about 13
covariates which are anthropometric measurements (e.g., height, weight
and thicknes of skin and fat in different body locations). The
covariates are are highly correlated whichs causes additional
challenge in the model selection. In total, we have 251
observations. The goal is to find the model which is able to predict
the amount of fat sufficiently well while requiring least amount of
measurements for a new person.

\cite{paper:bodyfat} report results using steplm with a significance
level of 0.157 with AIC selection \citep{akaike1974new}, fixing
abdomen and height to be always included in the model. For better
comparison we don't fix any of the variables.  The steplm approach is
carried out combining the \texttt{step} and \texttt{lm} functions

For the selection via projpred, the Bayesian reference model includes
all the covariates using a regularised horseshoe prior on the
covariate coefficients \citep{paper:rhs}. Submodels are explored using
forward search (the results are not sensitive to whether forward or
backward search is used), and the predictive utility is the expected
log-predictive density (elpd) estimated using PSIS-LOO
\citep{paper:psis_loo}.  We select the submodel as the one with the
smallest size which has an elpd score higher (i.e., better) than the
reference model score with probability of at least 5\%. See Appendix A
for brief review of projection predictive approach and papers by
\citet{paper:model_selection} and \citet{paper:projpred} for more
details. projpred selection is made using the \texttt{projpred} package.

The inclusion frequencies of each feature in the final model given 100
bootstrap samples are depicted in Figure
\ref{fig:inclusion_frequencies}. The projpred approach has only two
variables, `abdomen' and `weight', with inclusion frequencies above
50\% (`abdomen' is the only one included always), the third most
included is `wrist' at 44\%, and the fourth is `height' at 35\%, while
steplm has seven features above 50\%. Such a higher variability and
lower stability of steplm can be observed also in the bootstrap model
selection frequencies reported in Table
\ref{tab:model_frequencies}. For example, the first five selected
models have a cumulative frequency of 76\% with projpred, but only of
14\% with steplm. In addition, the size of the selected model is much
smaller when using projpred.
\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/inc_prob.pdf}
  \caption{Bootstrap inclusion frequencies after 100 bootstrap samples.}
  \label{fig:inclusion_frequencies}
\end{figure}

First two rows of Table \ref{tab:model_performances} report the
predictive performances, in terms of cross-validated root mean square
error (RMSE), of the full model and the selected models using either
the projection predictive approach or the stepwise backward
elimination.  There is no difference in predictive performance of the
selected models by different approaches even if there is difference in
the selected features.

We repeat the experiment with a modified data set by adding 84
unrelated noisy features. Last two rows of Table
\ref{tab:model_performances} report cross-validated RMSE, the size of
the selected model and the number of selected noisy features using
either the projection predictive approach or the stepwise backward
elimination.  The projection predictive approach has similar
predictive performance and the same submodel size as in the first
experiment, whereas the stepwise regression has worse predictive
performance and selects a much larger submodel including a large
number of irrelevant features.

Both projpred and steplm compare a large number of models using either
forward or backward search, which can lead to selection induced
overfitting, but even with 100 covariates projpred is able to select a
submodel with similar performance as the full model. Although projpred
uses Bayesian inference and steplm maximum likelihood, it is not
decisive feature explaining the difference in the performance.  The
main difference between the approaches is whether a reference model is
used or not, and in the rest of the paper we look into more detail how
the use of reference model helps and limitations of the approach.

\subsection{Benefits and cost of using a reference model}

A properly designed reference model is able to filter parts of the
noise present in the data, and hence provide an improved and more
stable selection process. This holds even if the reference model does
not perfectly resemble the true data generating process. Moreover, our
analyses indicate that the substantial reduction of variance
attributable to noise is usually more important than small potential
bias due to model misspecification.  We argue that, regardless of how
the reference model is set up and used in the inference procedure, it
can be always seen as acting as a filter on the observed
data. Furthermore, regardless of what specific model selection method
is used, a reference model can be used instead of raw data during the
selection process to improve the stability and selection
performance. Our results indicate that the core reason why the
reference model based methods perform well is the reference model
itself, rather than the specific way of using it.  In general, the
less data we have and the more complex the estimated models, the
higher the benefit of using a reference models as the basis for
variable selection.

If one of the models to be compared is the full model which can be
used as a reference model, there is no additional cost of using a
reference model. Sometimes including all the available features in an
elaborate model can often be computationally demanding.  In such a
case, even simpler screening or dimensionality reduction techniques,
as for example the supervised principal components
\citep{bair2006prediction,piironen2018} can produce useful reference
models.

The remainder of the paper is structured as follows. In Section
\ref{reference-model-approach}, we review the concept of the reference
model, its benefits with examples and how it can be used as a filter
on data in a simple way. In Section \ref{comparison}, we show the
benefits of a reference model approach, regardless of what specific
procedure is applied, using the normal means estimation problem as a
benchmark. Finally, we end with a conclusion in Section
\ref{conclusion}.

@@

A good predictive model is able to filter part of the noise present in
the data. The noise is the main source of the instability and tends to
obscure the relevance of the covariates in relation to the target
variable of interest. Let us consider the following simple
explanatory example taken from \cite{paper:projpred}. The data
generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim N(0,1) && \nonumber \\ 
     Y|&f\sim N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} N(\sqrt{\rho}f,1-\rho) \quad &&j=1,..,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} N(0,1) &&j=k+1,..,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy
observation. The first $k$ covariates are strongly related to the
target variable $Y$ and correlated among themselves. Precisely, $\rho$
is the correlation among any pair of the first $k$ covariates, whereas
$\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between
any relevant covariate and, respectively, $f$ and $Y$. If we had an
infinite amount of observations, the sample correlation would be equal
to the true correlation between $X_j$ and $Y$. However, even in this
ideal asymptotic regime, this correlation would still remain a biased
indicator of the true relevance of each covariate (represented by the
correlation between $X_j$ and $f$) due to the intrinsic noisy nature
of $Y$.

@@

@@

In this section, we analyse in more detail the benefits of reference
model approaches in combination of a few common variable selection
approach.

\subsection{Simulated data}

The normal means problem consists of estimating the (usually
sparse) vector of means of a vector of normally distributed
observations. The dimensionality of the vector of means is denoted by
$p$ and $\{z_{j}\}_{j=1}^{p}$ is the vector of observations of the
random variables $\{Z_{j}\}_{j=1}^{p}$. The task is to estimate the
latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model: \
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}), \quad j=1,..,p.
\end{equation}
This is equivalent to a linear regression where the design matrix is
the identity matrix with the number of features being equal to the
number of observations. This formulation can be found in practice, for
example, in the analysis of microarray data, where a large set of
genes are tested in two groups of patients labeled as positive or
negative to some disease \citep{paper:efron,efron2012large}. The
objective of the analysis is to select the subset of genes
statistically relevant to the disease. One common way to proceed is to
compute the two-sample $t$-statistic for every gene separately.  After
normalising these statistics, they then become the data, $Z$, in the
normal means problem \eqref{eq:normal_means_problem}. For further
details see the examples in \cite{paper:efron, efron2012large}.

In our experiments, we retrieve the normal means problem from the
sample correlations between the target and the covariates using the
Fisher $z$-transformation \citep{hawkins1989using}. Suppose we have a
continuous target random variable $Y$, a set of $p$ continuous
covariates $\{X_{j}\}_{j=1}^{p}$, and denote
$\rho_{j}=Cor(Y,X_{j})$. Further, suppose we have observed $n$
statistical units and define $r_{j}$ the sample correlation between
the observations of the target variables $\{y_{i}\}_{i=1}^{n}$ and the
$j$-th covariate $\{x_{ij}\}_{i=1}^{n}$. Finally, we refer to the
Fisher $z$-transformation function $\text{tanh}^{-1}(\cdot)$ as
$T_{F}(\cdot)$. Assuming each pair $(Y,X_{j})$ to be bivariate
normally distributed, the corresponding transformed correlations are
approximately normally distributed with known variance: \
\begin{equation} \label{eq:fisher_transformation}
T_{F}(r_{j})\overset{ind}{\sim} N(T_{F}(\rho_{j}),\frac{1}{n-3}) \quad j=1,..,p.
\end{equation}
Therefore, rescaling the quantities $T_{F}(r_{j})$ by $\sqrt{n-3}$ and
denoting the results as $z_{j}$, we have the formulation
\eqref{eq:normal_means_problem} of the normal means problem, this time
with unit variance: \
\begin{equation} \label{eq:normal_means_problem2}
Z_{j}|\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \quad j=1,..,p.
\end{equation}
Now the quantities of interest $\theta_{j}$ are equal to
$\sqrt{n-3}\,T_{F}(\rho_{j})$.

We use different levels of correlation $\rho\in\{0.3,0.5\}$ and number
observation $n\in\{50,70,100\}$. The total number of covariates $p$
and the number of relevant covariates $k$ are fixed to $p = 1000$ and
$k = 100$, respectively. In general, the lower $\rho$ and $n$, the
more challenging the selection.

\subsection{Model selection approach to be compared}

The Bayesian estimation methods we use to solve the normal means
problem have in common that first a Bayesian regression model is
fitted using some sparsifying prior on the regression coefficients,
for example, as discussed in \cite{paper:dirichlet_laplace},
\cite{bhadra2017horseshoe+} and \cite{johnstone2004needles}.  In our
experiments, we include the Dirichlet-Laplace
\cite[DL;][]{paper:dirichlet_laplace} prior and the regularised
horseshoe (RHS) prior \citep{paper:rhs}, which we think scales better
to higher number of covariates as compared to other horseshoe-like
priors.

% If these priors have continuous support, as is the case for the DL and
% horseshoe-like priors \citep{paper:hs,bhadra2017horseshoe+,paper:rhs},
% there is no canonical way to complete the selection since the
% posterior probability of an exactly zero signal is always
% zero. Whereas \cite{paper:dirichlet_laplace} used k-means clustering
% over the posterior samples to perform the actual selection,
% \cite{bhadra2017horseshoe+} just focused their analysis on the
% obtained shrinkage and the correct estimation of the sparse signals
% without any actual selection.

% For the Bayesian regression
% models with these priors, we compare the posterior shrinkage and the
% correct estimation of the sparse signals with and without using a
% reference model as a noise filter. For this purpose, we consider the
% average SSE (sum of square errors) and the average SESE (sum of
% expected square errors) as quantities measuring estimation
% accuracy. They are defined as \
% \begin{align}
% \text{SSE}&=\frac{1}{n-3}\sum_{j=1}^{p}(\hat{\theta}_{j} - \theta^{0}_{j})^{2} \label{eq:SSE} \\
% \text{SESE}&=\frac{1}{n-3}\sum_{j=1}^{p}\int_{\mathbb{R}}(\theta_{j}-\theta^{0}_{j})^{2}p(\theta_{j}|z_{1},..,z_{p})d\theta_{j} \\
% &=\frac{1}{n-3}\sum_{j=1}^{p}\mathbb{E}_{\theta_{j}}[(\theta_{j}-\theta^{0}_{j})^{2}|z_{1},..,z_{p}], \label{eq:SESE}
% \end{align}
% where $\theta_{j}^{0}=\sqrt{n-3}\,T_{F}(\rho_{j})$ is the true value,
% $\hat{\theta}_{j}$ is the posterior median of each parameter
% $\theta_{j}$. We divide the errors by $(n-3)$ to have quantities
% independent of the number of statistical units.  The SESE takes the
% posterior uncertainty in account and thus also the shrinkage of the
% posterior distributions, whereas the SSE only measures the bias in the
% point estimates.  Any (sensible) point estimate could be used to
% compute the SSE but here we follow \cite{paper:dirichlet_laplace} and
% use the posterior median. The results that we show later on in Section
% \ref{shrinkage-signal} are not sensitive to the choice of this point
% estimate.

% For shrinkage priors, one central aspect is the amount of shrinkage
% they actually impose on the relevant and irrelevant features. Ideally,
% the irrelevant features should be completely shrunk to zero while the
% relevant features should not be shrunk at all. The amount of shrinkage
% can be quantified by the shrinkage factor $\kappa_{j}$ for each
% parameter $\theta_{j}$ ranging from $0$ (no shrinkage) to $1$
% (complete shrinkage). For model \eqref{eq:normal_means_problem2} with
% a hierarchical shrinkage prior of the type
% \begin{align}
% Z_{j}|&\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \qquad j=1,..,p \\
% &\theta_{j}|\tau,\nu_{j}\overset{ind}{\sim}N(0,\nu_{j}^{2}\tau^{2})
% \end{align}
% the shrinkage factor $\kappa_{j}$ is defined as
% \begin{equation}
% \kappa_{j}=\frac{1}{1+\tau^{2}\nu_{j}^{2}}.
% \end{equation}
% \
% For RHS and DL priors we have
% \begin{equation}
% \kappa_{j}^{\text{RHS}}=\frac{1}{1+\tau^{2}\lambda_{j}^{2}}, \qquad \kappa_{j}^{\text{DL}}=\frac{1}{1+\psi\phi_{j}^{2}\tau^{2}}.
% \end{equation}
% In the left hand-side expression, $\tau$ and $\lambda_{j}$ represent
% the global and local shrinkage parameters of the regularised horseshoe
% prior \cite[for further details see][]{paper:rhs}, whereas on the
% right hand-side, $\psi\phi_{j}^{2}\tau^{2}$ is the variance term of
% the Dirichlet-Laplace prior of parameter $\theta_{j}$ following the
% notation of \cite{paper:dirichlet_laplace}. We use the posterior mean
% of $\kappa_{j}$ as a point estimate.
